### 2016-07-10 version0.0/intro0.0.tex
Update intro0.0.tex
edits to "Interstellar Medium" before first subsection

\section{Interstellar Medium}
\label{sec: ism_intro}
%how do we observe it
%what the observation tell us
% how do we measure the dust and gas
ISM in galaxies is a complex system with wide range of structures, but this structures can be divided into two main components: Gas and Dust.
Gas in ISM of galaxies is mostly made of hydrogen and helium with 1 per cent of heavier elements (metals), and can be in the forms of clouds or diffuse regions.
Gas clouds contains molecular clouds, cold atomic neutral hydrogen (\hi) or hot \hi~clouds, and diffuse gas can be either \hii~(ionized hydrogen) regions or hot intercloud mediums.
Dust mostly contains silicate, graphite grains can be seen in dark clouds or galactic cirrus.
%Among these regions in ISM, molecular clouds, \hii~regions and dust clouds have important role in understanding star formation in galaxies. 
Emission lines from atoms or ions (i.e \hii~regions) in a hot gas, lines from molecules in cold clouds and 21~cm emission of \hi~are the main observing signatures of the gas in the ISM of galaxies.
Dust can be observed directly through FIR and sub-mm observation or indirectly through their extinction and attenuation effect (see Sec.~\ref{sec: extinction}).

Molecular clouds in ISM are cold ($\sim10$ K) and dense ($ 50<n<500$ cm$^{-3}$), masses in molecular clouds range from few to million solar masses. %~\citep{Bolato08}.
Density in the molecular clouds is not uniform; dense and colds clumps.
Since these clumps are ideal places to form stars, studying molecular clouds has an important role in understanding formation of stars
However, H$_2$ molecules, the main component of molecular clouds, have no permanent electric dipole moment which makes them really hard to detect in visible light.
The second dominant component, helium, is a mono-atomic gas and has the same problem as hydrogen, but these clouds also contain heavier elements such as carbon monoxide, hydrogen cyanide, water, ammonia and more than tens of others.
CO molecules take a third place in case of abundance in the molecular clouds.
In fact, most of our knowledge from extragalactic molecular clouds comes from observing emission line of CO molecules, which due to their heavy weight, they have very low rotational state and can be excited in the temperature of molecular clouds. 
Other heavier molecules have emission lines which are detectable in spectrum of clouds (specially galactic ones), but observing CO emission is the most common way to study properties of molecular clouds i.e. their mass (see Sec.~\ref{sec: ismmap} for more details).

The molecular clouds are surrounded by cloud of atomic gas~\citep{Kennicutt12}, which are easily traceable by 21-cm emission.
Hydrogen atoms in the ISM (excluding the regions close to type ``O" and ``B" stars) are in their ground state. 
The electron and proton spins in the ground state can be parallel (i.e. both spin up) or anti-parallel (i.e. proton's spin is up and electron's spin is down or vice versa). 
The energy state of anti-parallel mode is slightly less than parallel mode.
Since atoms always want to be in the state with lowest energy possible, electrons in the parallel mode have a tendency to flip to the anti-parallel mode. 
However, difference between these two states are very small and it takes millions of years before the transition happens.
The large amount of hydrogen atoms, on the other hand, compensate the rareness of the transition and at any given time, there are enough hydrogen atoms to emit the 21-cm radiation. 
The observation of the \hi atoms is specifically necessary for understanding physical properties and dynamics of ISM, which leads to understanding star formation process in ISM~\citep{Walter08}.

\hii~regions are hot and low density regions that are mostly located in the disk of spiral galaxies.
Optically bright \halpha emission from \hii~regions correlates with number of new massive stars in regions and can be used as star formation tracers~\citep[e.g.][]{Kennicutt98b,Calzetti13}. 
High energy UV photon emitted by massive new born stars have sufficient amount of energy to ionize their surrounding gas.
~\halpha emission is main observational feature\hii~regions.
Temperature and density of these regions can be estimated by observing other recombination lines such as H$\beta$, \sii, \oii, \oiii, and \nii from \hii~regions.
Studying \hii~regions provide valuable insight into star forming regions by providing information such as star formation rate, initial mass function (IMF), and distribution of ionizing stellar masses~\citep[][and references therein]{Azimlu11}.

Dust has insignificant effect on mass of ISM, but it plays crucial roles in evolution of ISM.
The total gas mass of galaxies can be determined using optical depth in sub-mm waveband of dust emission. 
In spiral galaxies dust are in cold regions $\sim 30$~K and has power low distribution of size.
Source of dust thermal emission is from stars and shape of the continuum depends on size of the grain.
MIR emission from dust grains are manly dominated by PAHs heated by evolved stellar population~\cite{Smith07a} or recently formed types``A" and ``B'' stars~\cite{Peeters04}.
Large grains that are either heated by light from star forming regions or by interstellar radiation field from total stellar population emit most of their light in FIR~\citep[e.g.][]{Calapa14, lu14}
Dust are main source of extinction and reddening of star lights (See Sec.\ref{sec: extinction} for more information).

>>>>

\section{Interstellar Medium} %PB: suggest "Components of the Interstellar Medium in Galaxies" or "The Interstellar Medium in Galaxies"
\label{sec: ism_intro}
%how do we observe it
%what the observation tell us
% how do we measure the dust and gas
The interstellar medium in galaxies is a complex system with a wide range of structures, but these structures can be divided into two main components: gas and dust. %PB: in general, don't start a sentence with an acronym.
The gas in the ISM of galaxies is mostly made of hydrogen and helium with 1 per cent heavier elements (metals), and can be in the form of clouds or diffuse regions.
Gas clouds include molecular clouds, cold neutral atomic hydrogen (\hi) or hot \hi~clouds, and diffuse gas can be either \hii~(ionized hydrogen) regions or hot intercloud medium.
Dust mostly contains silicate, graphite grains can be seen in dark clouds or galactic cirrus.
%Among these regions in ISM, molecular clouds, \hii~regions and dust clouds have important role in understanding star formation in galaxies. 
Emission lines from atoms or ions (i.e \hii~regions) in a hot gas, lines from molecules in cold clouds and 21~cm emission of \hi~are the main observational signatures of the gas in the ISM of galaxies.
Dust can be observed directly through far-infrared and sub-millimeter wavelength emission or indirectly through its extinction and attenuation effects (see Sec.~\ref{sec: extinction}).

Molecular clouds in the ISM are cold ($\sim10$ K) and dense ($ 50<n<500$ cm$^{-3}$); masses of molecular clouds range from a few to millions of solar masses. %~\citep{Bolato08}. %PB: why no citation?
Density in the molecular clouds is not uniform; dense and colds clumps. %PB: I don't get this sentence.
Since these clumps are ideal places to form stars, studying molecular clouds has an important role in understanding formation of stars
However, H$_2$ molecules, the main component of molecular clouds, have no permanent electric dipole moment which makes them difficult to detect in visible light. %PB: just visible light, or all wavelengths?
The second dominant component, helium, is a mono-atomic gas and has the same problem as hydrogen, but these clouds also contain heavier elements such as carbon monoxide, hydrogen cyanide, water, ammonia and more than tens of others.
Carbon monoxide (CO) molecules are the third-most abundant constituent of molecular clouds.
In fact, most of our knowledge from extragalactic molecular clouds comes from observing line emission of CO molecules, which {\bf due to their heavy weight, they have very low rotational state} and can be excited in the temperature of molecular clouds. %PB just extragalactic, or Galactic too? Phrase marked with \bf needs clarification
Other heavier molecules have emission lines which are also detectable in the spectrum of clouds (specially galactic ones), but observing CO emission is the most common way to study the properties of molecular clouds i.e. their mass (see Sec.~\ref{sec: ismmap} for more details).

The molecular clouds are surrounded by clouds of atomic gas~\citep{Kennicutt12}, which are easily traceable by 21-cm emission.
Hydrogen atoms in the ISM (excluding the regions close to hot stars, see below) are in their ground state. 
The electron and proton spins in the ground state can be parallel (i.e. both spin up) or anti-parallel (i.e. the proton's spin is up and the electron's spin is down or vice versa). 
The energy state of the anti-parallel mode is slightly less than the parallel mode.
Since atoms always want to be in the state with lowest energy possible, electrons in the parallel mode have a tendency to flip to the anti-parallel mode. 
However, the difference between these two states is very small and it takes millions of years before the transition happens.
The large amount of hydrogen atoms compensates for the rareness of the transition and at any given time, there are enough hydrogen atoms to emit the 21-cm radiation. 
Observation of the \hi atoms is specifically necessary for understanding physical properties and dynamics of the ISM, which leads to understanding the star formation process~\citep{Walter08}.

\hii~regions are hot and low density regions that are mostly located in the disks of spiral galaxies.
High energy UV photons emitted by massive new born stars have sufficient energy to ionize their surrounding gas, and
~\halpha emission is the main observational feature of \hii~regions.
The temperature and density of these regions can be estimated by observing other recombination lines such as H$\beta$, \sii, \oii, \oiii, and \nii. 
Optically bright \halpha emission from \hii~regions correlates with number of new massive stars in regions and can be used as a star formation tracer~\citep[e.g.][]{Kennicutt98b,Calzetti13}. %PB: moved this sentence.
Studying \hii~regions provides valuable insight into star forming regions by providing information such as star formation rate, initial mass function (IMF), and distribution of ionizing stellar masses~\citep[][and references therein]{Azimlu11}.

%PB: it is difficult to see how the sentences in this paragraph fit together. Right now it reads like a collection of barely-related facts.
Dust has an insignificant effect on the mass of the ISM, but it plays crucial roles in the ISM's evolution.
The total gas mass of galaxies can be determined using optical depth in sub-mm waveband of dust emission. 
In spiral galaxies dust grains are located in cold regions ($\sim 30$~K) and have a power-law distribution of size.
Source of dust thermal emission is from stars and shape of the continuum depends on size of the grain.
Mid-infrared emission from dust grains is mainly dominated by PAHs heated by evolved stellar population~\cite{Smith07a} or recently formed type ``A" and ``B'' stars~\cite{Peeters04}. %PB: A and B or O and B?
Large grains that are either heated by light from star-forming regions or by the interstellar radiation field from total stellar population emit most of their light in far-infrared~\citep[e.g.][]{Calapa14, lu14}
Dust is the main source of extinction and reddening of starlight (See Sec.\ref{sec: extinction} for more information).
